WHITE WINE ANALYSIS  by Jose Maria G.M
========================================================

We will analyze the variables of the white wine dataset, the variables, their relationships and try to determine what determines a high quality white wine.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.


#This sets the global heights and widths for the plots in the project.
#This approach makes your plot sizes uniform, page-width and centered.

knitr::opts_chunk$set(fig.width=9,fig.height=5,fig.path='Figs/',
                      fig.align='center',tidy=TRUE,
                      echo=FALSE,warning=FALSE,message=FALSE)

# Load all of the packages that you end up using in your analysis in this code
# chunk.
library(ggplot2)
library(dplyr)
library(GGally)
library(knitr)
```

```{r echo=FALSE, Load_the_Data}

# Load the Data 
wine <- read.csv("wineQualityWhites.csv")
```

# Univariate Plots Section

We are going to perform some preliminary exploration of the dataset. Let's determine what are the fields of the dataset.

```{r echo=FALSE, message=FALSE, warning=FALSE}

#show all variables
names(wine)
```

Which are the number of observations and the kind of variables.

```{r echo=FALSE, Univariate_Plots}
#show kind of data in each variable
str(wine)
```

Summary of each variable.
```{r echo=FALSE, message=FALSE, warning=FALSE}

# statistical data of variables
summary(wine)
```

We can see that in the sample there are more wine of quality between 5 and 7. It would nice a dataset with more observations about quality 9, we only have 5 observations.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# table with number of observations per quality values
table(wine$quality)
```
 
# Univariate Analysis
Now we are going to plot the histogram of the variables of the dataset white wine.


Plot citric acid:<br>
It has normal distribution skewed to the right with some outliers.
```{r echo=FALSE, message=FALSE, warning=FALSE}

# Histogram -  Citric Acid
ggplot(aes(x = citric.acid), data = wine) + 
            geom_histogram(binwidth = 0.01)

```

Plot residual sugar:<br>
The distribution has very high values for low values of residual sugar.
```{r echo=FALSE, message=FALSE, warning=FALSE}

# Histogram -  Residual sugar
ggplot(aes(x = residual.sugar), data = wine) +  
            geom_histogram(binwidth = 0.1)
```


Plot chlorides:<br>
It has a normal distribution with a hight number of outliers.

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Histogram -  Chlorides
ggplot(aes(x = chlorides), data = wine) +  
            geom_histogram(binwidth = 0.001)
```

Plot Free Sulfur Dioxides:<br>
It has a normal distribution skewed to the right.

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Histogram -  Free Sulfur Dioxide
ggplot(aes(x = free.sulfur.dioxide), data = wine) +  
            geom_histogram(binwidth = 1)
```
Plot Total Sulfur Dioxides:<br>
It has a normal distribution with some outliers.

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Histogram -  Total Sulfur Dioxide
ggplot(aes(x = total.sulfur.dioxide), data = wine) +   
            geom_histogram(binwidth = 1)
```
Plot Density:<br>
It has a normal distribution with values between 0.99 and 1.0. with some outliers.

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Histogram -  Density
ggplot(aes(x = density), data = wine) +   
            geom_histogram(binwidth = 0.0001)
```


Plot pH:<br>
It has a normal distribution with a mean of 3.188, it is acid and this gives a refreshing sensation.
```{r echo=FALSE, message=FALSE, warning=FALSE}

# Histogram -  pH
ggplot(aes(x =  pH ), data = wine) +   
            geom_histogram(binwidth = 0.01)
```


Plot Sulphates:<br>
It has a normal distribution skewed to the right.

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Histogram -  Sulphates
ggplot(aes(x =  sulphates ), data = wine) +   
            geom_histogram(binwidth = 0.01)
```


Plot Alcohol:<br>
It has not a normal distribution with values between 8 and 14.
```{r echo=FALSE, message=FALSE, warning=FALSE}

# Histogram -  Alcohol
ggplot(aes(x =alcohol), data = wine) +   
            geom_histogram(binwidth = 0.1)
```


Plot Quality:<br>
It has a normal distribution.
```{r echo=FALSE, message=FALSE, warning=FALSE}

# Histogram -  Quality
ggplot(aes(x =quality), data = wine) +   
            geom_histogram(binwidth = 1)
```

Plot Volatile Acidity:<br>
It has a normal distribution skewed to the right.

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Histogram -  Volatile acidity
ggplot(aes(x = volatile.acidity), data = wine) +
             geom_histogram(binwidth = 0.01)
```

Plot Fix Acidity:<br>
It has a normal distribution.

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Histogram -  Fixed acidity
ggplot(aes(x = fixed.acidity), data = wine) +
              geom_histogram(binwidth = 0.1)

```
 
# Bivariate Plots Section

We are going to create a matrix where we can see if there is some correlation between variables.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=9, fig.height=7 }

theme_set(theme_minimal(12))

#set the SEED for reproducible results
set.seed(1836)
#create a new dataset without variable "x"
wine_subset <- subset(wine, select = -c(X))
 
#matrix plot with correlations
ggpairs(wine_subset)+
  theme(text = element_text(size=8), 
     axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size=7))
 
```
Here we see that there are some variables with a high correlation:<br>
   - density and alcohol  -0.78
   - free.sulfur.dioxide and total.sulfur.dioxide  0.616
   - residual.sugar and density  0.839
   - quality and alcohol   0.436
   - density - total.sulfur.dioxide  0.53

First examine some relation of variables with low correlation


We see that all white wines in the sample have pH values between 2.72 and 3.82 and the mean value increase for higher quality.
```{r echo=FALSE, Bivariate_Plots}

#plot quality - pH 
ggplot(aes(x = quality, y = pH), data = wine)  +
       geom_jitter(alpha=0.6, width = 0.25, height = 0) +
       geom_line(stat = 'summary', fun.y = mean, color = "blue")

```


The residual sugar mean does not have a great variations for the quality values.
```{r echo=FALSE, message=FALSE, warning=FALSE}

#plot quality - residual.sugar 
ggplot(aes(x = quality, y = residual.sugar), data = wine) +
           geom_jitter(alpha=0.6, width = 0.25, height = 0) +
           geom_line(stat = 'summary', fun.y = mean, color = "blue")
      
```

Plot alcohol-quality:<br>
There is a correlation of 0.436, and we see how the quality increases with alcohol mean.
```{r echo=FALSE, message=FALSE, warning=FALSE}

#plot quality - alcohol
ggplot(aes(x = quality, y = alcohol), data = wine)  +
           geom_jitter(alpha=0.6, width = 0.25, height = 0) + 
           geom_line(stat = 'summary', fun.y = mean, color = "blue")
 

```


Plot alcohol-quality:<br>
Here we see some outliers and when alcohol descreases incresases the density.

```{r echo=FALSE, message=FALSE, warning=FALSE}

#plot density - alcohol 
ggplot(aes(x = density, y = alcohol), data = wine)  + 
           geom_point(alpha = 1/20) + 
           geom_smooth()
        
```

 
    
Plot free.sulfur.dioxide - total.sulfur.dioxide:<br>
There is a correlation 0.616 and there are some outliers. Here when free.sulfur.dioxide increases the total.sulfur.dioxide increases.    
```{r echo=FALSE, message=FALSE, warning=FALSE}

#plot free.sulfur.dioxide - total.sulfur.dioxide
ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide), data = wine)  + 
           geom_point(alpha = 1/15) + 
           geom_smooth()

```
    
Plot density - total.sulfur.dioxide:<br>
There is a correlation 0.53 and there are some outliers. Here when density increases the total.sulfur.dioxide increases.       
    
```{r echo=FALSE, message=FALSE, warning=FALSE}

#plot density - total.sulfur.dioxide
ggplot(aes(x = density, y = total.sulfur.dioxide), data = wine)  + 
           geom_point(shape = 1, alpha = 1/10) + 
           geom_smooth()

```
  
        
```










 

```{r echo=FALSE, message=FALSE, warning=FALSE}
#create quality groups
quality_groups <- group_by(wine, quality)
#calculate for each group the differents variables means.
groups.wine <- summarise(quality_groups,
          alcohol_mean = mean(alcohol),
          residual.sugar_mean = mean(residual.sugar),
          density_mean = mean(density),
          pH_mean = mean(pH),
          fixed.acidity_mean = mean(fixed.acidity),
          volatile.acidity_mean = mean(volatile.acidity),
          citric.acid_mean = mean(citric.acid),
          sulphates_mean = mean(sulphates), 
          chlorides_mean = mean(chlorides),
          free.sulfur.dioxide_mean = mean(free.sulfur.dioxide),
          total.sulfur.dioxide_mean = mean(total.sulfur.dioxide),
          n = n())

#arrage by age
groups.wine <- arrange(groups.wine, quality)
 
```

# Bivariate Analysis

We have seen that there are some strong relation between variables:<br>
    - density and alcohol   
    - free.sulfur.dioxide and total.sulfur.dioxide   
    - residual.sugar and density   
    - quality and alcohol 
    - density and toal sulfur dioxide

Here we have the means of the variables of the dataset of white wine.

```{r echo=FALSE, message=FALSE, warning=FALSE}

#create table1 
table1 <- groups.wine

#remove columns 
table1 <- table1[,-(8:12)]

#show table
kable(table1,caption = "Summary means part 1")


```

```{r echo=FALSE, message=FALSE, warning=FALSE}

#create table2 
table2 <- groups.wine

#remove columns 
table2 <- table2[,-(2:7)] 

#show table
kable(table2,caption = "Summary means part 2")

```
 
  

# Multivariate Plots Section

Plot residual sugar and density: here we see how increases density when increases residual sugar.
```{r echo=FALSE, message=FALSE, warning=FALSE}

#plot residual.sugar and density
ggplot(aes(x = residual.sugar, y = density), data = wine)  + 
           geom_point(alpha = 1/30, color = "blue") + 
           geom_smooth(method = "lm", color= 'red',size = 1) 
```


Plot density and chlorides: here we see how decreases chlorides and density increases quality.

```{r echo=FALSE, message=FALSE, warning=FALSE}

#create subset wine
wine_multi <- subset(wine, density <quantile(wine$density,0.95) &
                       chlorides < quantile(wine$chlorides,0.95))
#plot density and chlorides with mean values                       
ggplot(aes(x = density, y = chlorides), data = wine_multi) +
  geom_point(alpha = 0.8, aes(color = as.factor(quality))) +
  geom_point(aes(x = density_mean, y = chlorides_mean, size = quality,
                 fill=as.factor(quality)), data = groups.wine, pch = 21)
```

Plot free.sulfur.dioxide and total.sulfur.dioxide: we can see if free.sulfur.dioxide increases also increases total.sulfur.dioxide.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# plot free.sulfur.dioxide and total.sulfur.dioxide
ggplot(aes(x = total.sulfur.dioxide, y = free.sulfur.dioxide), data = wine_multi) +
      geom_point(alpha = 0.8, aes(color = as.factor(quality))) +
       
      geom_point(aes(x = density_mean, y = chlorides_mean, size = quality,
                 fill=as.factor(quality)), data = groups.wine, pch = 21) +
       
      scale_color_brewer(type = "div", palette = "RdYlBu", name="Quality",
                     direction=-1)+ # color the scatterplot by quality
      scale_fill_brewer(type = "div", palette = "RdYlBu", name="Quality",
                    direction=-1) + # color the mean points by Quality
      xlim(0,250) +
      ylim(0,150)

```


Plot density and alcohol: we can see if alcohol increases also decreases density and increases quality.
```{r echo=FALSE, message=FALSE, warning=FALSE}
#plot density and alcohol 
ggplot(aes(x = density, y = alcohol), data = wine)  + 
           geom_point(aes(color = as.factor(quality)), alpha= 0.9) +
           geom_smooth(color= 'red',size = 1) +
           xlim(0.985,1.0) +
           geom_point(aes(x = density_mean, y = alcohol_mean, size = quality,
                 fill=as.factor(quality)), data = groups.wine, pch = 21)
              
```

# Multivariate Analysis

We have seen how with high alcohol, low density we have a high quality wine.<br>
Also chlorides and density decreases, quality increases.


------

# Final Plots and Summary

### Plot One


```{r echo=FALSE, Plot_One}

#plot residual.sugar and density
ggplot(aes(x = residual.sugar, y = density), data = wine)  +
  geom_point(aes(color = as.factor(quality)), alpha= 0.8) +  # color the points by quality level
  xlab("residual sugar") +
  geom_smooth(method = "lm", color= 'red',size = 1) +
  xlim(0,20)  +
  geom_point(aes(x = residual.sugar_mean, y = density_mean, size = quality,
                 fill=as.factor(quality)), data = groups.wine, pch = 21) +
  scale_color_brewer(type = "div", palette = "RdYlBu", name="Quality",
                     direction=-1)+ # color the scatterplot by quality
  scale_fill_brewer(type = "div", palette = "RdYlBu", name="Quality",
                    direction=-1) + # color the mean points by Quality  
  ylab(expression(Density~(g/cm^{3}))) +
  xlab(expression(Residual~Sugar~(g/dm^{3})))  +
  ggtitle("Alcohol and Residual Sugar, by Quality")
 
```

### Description One
In the previous plot we see density and residual sugar with a correlation of 0.839 is the greater value in the whole variables. It can use a linea regression to estimate the values. If residual sugar decrease density and increases the quality of the white wine.

### Plot Two
```{r echo=FALSE, Plot_Two}

#plot density and alcohol 
ggplot(aes(x = density, y = alcohol), data = wine)  + 
           geom_point(aes(color = as.factor(quality)), alpha= 0.9) +
           xlab(expression(Density~(g/cm^{3}))) +
           ylab(expression(Alcohol~("% by volume"))) +
           geom_smooth(color= 'red',size = 1) +
           xlim(0.985,1.0) +
           geom_point(aes(x = density_mean, y = alcohol_mean, size = quality,
                 fill=as.factor(quality)), data = groups.wine, pch = 21) +
           scale_size_continuous(range = c(3, 10)) + # change the default size option
           scale_color_brewer(type = "div", palette = "RdYlBu", name="Quality",
                     direction=-1)+ # color the scatterplot by quality
          scale_fill_brewer(type = "div", palette = "RdYlBu", name="Quality",
                    direction=-1) + # color the mean points by Quality
          guides(size=FALSE) + # don't include the size legend
           ggtitle("Density and Alcohol and mean")
```

### Description Two
In the previous plot we see the high correlation 0.78 between alcohol and density.When density decreases then alcohol increases and the quality of the wine.

### Plot Three

```{r}
# plot density and chloides with means
ggplot(aes(x = density, y = chlorides), data = wine_multi) +
      geom_point(alpha = 0.8, aes(color = as.factor(quality))) +
      xlab(expression(Density~(g/cm^{3}))) +
      ylab(expression("Sodium chloride"~(g / dm^{3}))) +
      geom_point(aes(x = density_mean, y = chlorides_mean, size = quality,
                 fill=as.factor(quality)), data = groups.wine, pch = 21) +
      geom_smooth(color= 'red',size = 1) +
      scale_size_continuous(range = c(3, 10)) + # change the default size option
      scale_color_brewer(type = "div", palette = "RdYlBu", name="Quality",
                     direction=-1)+ # color the scatterplot by quality
      scale_fill_brewer(type = "div", palette = "RdYlBu", name="Quality",
                    direction=-1) + # color the mean points by Quality
      guides(size=FALSE) +# don't include the size legend
           ggtitle("Density and Chlorides, by quality")   
```



### Description Three
In this plot we see how with low density and low chlorides we have high quality white wines.

 

# Reflection

We have some variables with high correlations:<br>
    - density and alcohol  
    - free.sulfur.dioxide and total.sulfur.dioxide  
    - residual.sugar and density   
    - quality and alcohol 
    
Also we have discovered that we high alcohol  and low density we have high quality white wines. And with low chlorides and low density we have high quality white wines.

Here we have a few variables and I only have checked linear correlation, which would be the methology for a high number of variables? how could I detect other kind of relation between two or more variables?

In relation to R I have had some problems to search a way to show the tables of means variables.  
  
With these infomation now we could predict the quality of a white wine.
 